Strong lensing selection effects

### 2023-07-13 paper/seleff.tex
resubmitted version

  \abstract
  % context heading (optional)
  % {} leave it empty if necessary  
    {
Strong lenses are a biased subset of the general population of galaxies.
}
  % aims heading (mandatory)
   {
The goal of this work is to quantify how lens galaxies and lensed sources differ from their parent distribution, namely the strong lensing bias.
} 
   % methods heading (mandatory)
   {
We first studied how the strong lensing cross-section varies as a function of lens and source properties. 
Then, we simulated strong lensing surveys with data similar to that expected for Euclid and measured the strong lensing bias in different scenarios.
We focused particularly on two quantities: the stellar population synthesis mismatch parameter, $\asps$, defined as the ratio between the true stellar mass of a galaxy and the stellar mass obtained from photometry, and the central dark matter mass at fixed stellar mass and size.
}
% results heading (mandatory)
   {
Strong lens galaxies are biased towards larger stellar masses, smaller half-mass radii and larger dark matter masses.
The amplitude of the bias depends on the intrinsic scatter in the mass-related parameters of the galaxy population and on the completeness in Einstein radius of the lens sample.
For values of the scatter that are consistent with observed scaling relations and a minimum detectable Einstein radius of $0.5''$, the strong lensing bias in $\asps$ is $10\%$, while that in the central dark matter mass is $5\%$.
The bias has little dependence on the properties of the source population: samples of galaxy-galaxy lenses and galaxy-quasar lenses that probe the same Einstein radius distribution are biased in a very similar way.
%Quadruply imaged quasar lenses, however, are biased towards higher ellipticity galaxies.
%{\bf
%The bias on time-delay measurements of $H_0$ due to the use of prior information from galaxy-galaxy lenses is on the order of $1\%$.
%}
}
  % conclusions heading (optional), leave it empty if necessary 
   {
Given current uncertainties, strong lensing observations can be used directly to improve our current knowledge of the inner structure of galaxies, without the need to correct for selection effects.
{\bf Time-delay measurements of $H_0$ from lensed quasars can take advantage of prior information obtained from galaxy-galaxy lenses with similar Einstein radius.}
%{\bf Galaxy-galaxy lenses can be used to inform the properties of quasar lenses for the purpose of measuring $H_0$ from time delays
}
   \keywords{
             Gravitational lensing: strong 
               }

@@

{\bf
Strong lenses, however, are a biased subset of the general population of galaxies and background sources.
In order for a lens-source pair to be included in a strong lensing survey, the following conditions must be met:
\begin{enumerate}
\item The lens must produce multiple images of the source
\item The multiple images must be detected and resolved
\item The system must be recognized as a lens by the survey.
\end{enumerate}
Condition 1 poses constraints on the mass distribution of the lens, as well as on the geometry of the system. 
First of all, not all lenses are capable of producing multiple images, but only those whose surface mass density $\Sigma(\boldsymbol\theta)$ is larger than the critical surface mass density for lensing $\Sigma_{\mathrm{cr}}$ at at least one position $\boldsymbol\theta$ \citep{SEF92}.
This condition excludes objects with a very diffuse mass distribution from the population of lenses.
Second, objects with a higher concentration of mass allow for a larger portion of the source plane to be multiply imaged, and therefore are over-represented in lens samples. Third, at fixed lens properties, sources that are located further away along the line of sight are more likely to be lensed.

%First, the projected surface mass density $\Sigma(\boldsymbol\theta)$ must be larger than the critical surface mass density for lensing $\Sigma_{\mathrm{cr}}$ at at least one position $\boldsymbol\theta$ \citep{SEF92}, a condition that excludes objects with a very diffuse mass distribution from the population of lenses.
%Second, the source must be sufficiently close in projection to the lens center, in a way that depends on the lens mass distribution. Galaxies with a higher concentration of mass allow for a larger portion of the source plane to be multiply imaged, and therefore are over-represented in lens samples. Third, at fixed lens properties, sources that

@@

%A necessary condition for a galaxy to act as a strong lens with respect to a given source is that its projected surface mass density $\Sigma(\boldsymbol\theta)$ must be larger than the critical surface mass density for lensing $\Sigma_{\mathrm{cr}}$ at at least one position $\boldsymbol\theta$ \citep{SEF92}.
%This condition excludes objects with a very diffuse mass distribution from the population of lenses.
%In general, galaxies with a higher concentration of mass are more likely to be strong lenses, and are therefore over-represented in lens samples.
%The distribution of strongly lensed sources can also be biased with respect to the general population of background galaxies: for instance, lensing magnification allows the detection of fainter objects with respect to the field.

The cumulative effect of these conditions can be summarised as follows.
}
%In general, 
The probability distribution $\prsl$ of a sample of strong lenses from a given survey with selection criterion $S$ is given by \citep{Son22}:
\begin{equation}\label{eq:one}
\prsl(\psilens,\psisource|S) \propto \prlens(\psilens)\prsource(\psisource)\psel(\psilens,\psisource|S).
\end{equation}
In the equation above, $\psilens$ is the set of parameters describing the properties of foreground galaxies that are relevant for lensing, such as their redshift and mass distribution; $\psisource$ is the set of parameters describing background sources; $\prlens$ and $\prsource$ describe the parent distribution of foreground galaxies and background sources in the absence of lensing; and $\psel$ describes the probability of selecting a lens-source system with parameters $\psilens$ and $\psisource$ given the criterion $S$ used to define a lens.
This last factor takes into account both physical effects, that is whether a lens with parameters $\psilens$ produces a strongly lensed image of a source with parameters $\psisource$, and survey selection effects, that is whether such an image can be detected and recognised as a strong lens.

@@

{\bf Finally, we considered galaxy-quasar lenses. 
The strong lensing bias of lensed quasars was studied by \citep{C+C16}, in the context of power-law lens mass models with external convergence.
Here we addressed the question of how different are galaxy-quasar lenses from galaxy-galaxy lens samples. 
}
%Finally, we addressed the question of how different are galaxy-galaxy lens samples from sets of galaxy-quasar lenses. 
This %last% 
point is relevant for time-delay cosmography studies, in which measurements of the time-delay between the multiple images of a strongly lensed quasar are used to constrain the expansion rate of the Universe. Galaxy-galaxy lenses can in principle be used to help break some of the model degeneracies affecting these measurements \citep{Bir++20,B+T21}, but any difference between the two lens classes can introduce biases, if not corrected for.
With this study we aim to quantify this bias.

@@

%\subsubsection{Lensed galaxies}

@@

{\bf The first four rows of \Fref{fig:quasarbias} show} the strong lensing bias as a function of minimum Einstein radius in the following quantities: stellar population synthesis mismatch parameter, halo mass at fixed stellar mass, enclosed dark matter mass at fixed stellar mass and half-light radius, and axis ratio.
The biases of the population of lensed quasars (red curves) are very similar to that of lensed galaxies (blue curves), especially for values of $\theta_{\mathrm{Ein},min}<1''$, 
When considering only quad lenses, however, there are some differences, the most remarkable of which is the bias in the axis ratio: quad lenses tend to be on average galaxies with a higher ellipticity.
This is a well-known effect \citep[see e.g.][]{KKS97} which can be explained in the context of the results of section \ref{ssec:axisymmpoint}: the higher the ellipticity, the larger the area enclosed within the inner caustic, which is where a source needs to lie in order to produce four or more images.
%This trend was expected, given the results of section \ref{ssec:axisymmpoint}: the higher the ellipticity, the larger the area enclosed within the inner caustic, which is where a source needs to lie in order to produce four or more images.
Quad lenses tend to also have a slightly larger halo mass at fixed stellar mass, for $\theta_{\mathrm{Ein},min}<1''$.

@@

{\bf
A measurement of the time delay between the images of a lensed quasar can be converted into an estimate of the expansion rate of the Universe, $H_0$. %, provided that the mass distribution of the lens galaxy (and of matter along the line-of-sight) is accurately known.
In general, though, the value of $H_0$ is degenerate with the mass distribution of the lens (and of matter along the line-of-sight), which can be difficult to constrain on an individual lens basis.
One possible strategy to break this denegeracy is to use information gathered from a larger set of lenses as a prior for the time-delay lenses.
This was the strategy adopted by \citet{Bir++20} in the most recent measurement of $H_0$ by the TDCOSMO collaboration: they combined a larger sample of extended source lenses with a smaller set of time-delay lenses.
In this section we want to quantify the possible bias on $H_0$ that can result from this approach.

The lens mass model used by \citet{Bir++20} consists in a power-law density profile modified by a mass-sheet transformation. The corresponding projected density profile is
\begin{equation}\label{eq:mstprofile}
\kappa(\theta) \propto \lmst\theta^{1-\gamma} + (1 - \lmst),
\end{equation}
where $\lmst$ is called the mass-sheet transformation parameter.
Strong lensing data can constrain very well the power-law index $\gamma$, but not at all $\lmst$.
%, on which the inference on the Hubble constant depends.
Given the time delay, the inferred value of the Hubble constant $H_0$ scales linearly with $1/\lmst$.
\citet{Bir++20} measured the population distribution of $\lmst$ on the sample of extended source lenses, by combining strong lensing and stellar kinematics data, and then used it as a prior in the time-delay analysis.
Although the mass model used in our work is different from that of \Eref{eq:mstprofile}, we can estimate the bias associated with this approach by computing the value of $\lmst$ for each lens and then comparing the strong lensing bias on $\lmst$ of extended source and quasar lenses.
Appendix~\ref{sect:appendixc} explains how we obtained $\lmst$ for each lens. The strong lensing bias on the median value of $\lmst$ is shown in the fifth panel of \Fref{fig:quasarbias}\footnote{Since $\lmst$ is only defined for strong lenses, the median value of $\lmst$ of the general population of galaxies is not defined.}.

At fixed minimum Einstein radius, there is a small difference between the median $\lmst$ in extended source and quasar lenses, typically of order 1\%.
The associated bias on $H_0$ is, therefore, equally small.
The value of $\lmst$, however, increases with increasing minimum $\tein$. This implies that if the two samples probe different ranges in Einstein radius, the bias on $H_0$ can be larger than that.
The bias can also be larger if the properties of the general population of galaxies (as opposed to the strong lenses) are used to inform the structure of time-delay lenses \citep[see][]{C+C16}. This is simply because quasar lenses are on average more similar to galaxy-galaxy lenses than they are to the underlying galaxy population (the blue and red curves in \Fref{fig:quasarbias} are closer to each other than they are to the dashed horizontal lines).
}

@@

Strictly speaking, the results shown in \Sref{sect:results} apply only to samples of lenses and background sources with the same properties as our simulations.
{\bf Real lens samples can have more complex source distributions. For example, many existing lens-finding algorithms are tuned to identify sources with different colours from the lenses, introducing a selection that can modify the source distribution in redshift-magnitude space.
}
However, the comparison of section \ref{ssec:quasarbias} shows that, at fixed foreground galaxy population, the strong lensing bias for the sample of extended sources is indistinguishable from that of the lensed quasar population, at least for values of $\theta_{\mathrm{Ein,min}} < 1.0''$.
This result suggests that the dependence of the strong lensing bias on the details of the background source population is very weak, as long as no additional selection on the image configuration is applied (e.g. by selecting only quad lenses).
This is not a coincidence, but is a consequence of the fact that the dependence of the strong lensing cross-section on the lens parameters is a weak function of source magnitude\footnote{The cross-section itself is a strong function of source magnitude, but the trends between $\crosssect$ and the lens parameters are not.}, especially for magnitudes close to the detection limit (see section \ref{ssec:axisymmpoint}).
We can then conclude that the details of the properties of the background source population play a secondary role in determining the strong lensing bias.
{\bf 
A selection aimed at specific kinds of sources, such as blue star-forming galaxies or quasars, results primarily in different numbers of lenses being discovered, but does not change how biased the lenses are with respect to the general galaxy population.
}

@@

{\bf
The simulations described in \Sref{sect:lenspop}, on which our experiment is based, are examples of so-called lens-selected surveys:
the lens samples are built by first defining a population of possible lens galaxies and then searching for strong lenses among them. Photometric lens searches are typically carried out in this way \citep[see e.g.][]{Gav++12,Son++18a,Pet++19}.
An alternative approach is to build a source-selected sample, by first searching for signatures of strong lensing within a population of possible background sources and then obtaining information on the lens galaxies \citep[see e.g.][]{Hez++13}.
In principle, these two approaches lead to different strong lensing biases. %Therefore, strictly speaking, our results only apply to a lens-selected sample.
%For example, from \Fref{fig:lenspars} we can see that the distribution in redshift and stellar mass of the lenses extends beyond the bounds applied to the population of foreground galaxies: in principle, a source-based selection method 
For example, from \Fref{fig:lenspars} we can see that the stellar mass distribution of the lenses extends below the minimum mass cut that we applied to the population of foregrounds galaxies: a source-based search that is able detect those lenses would result in a different sample and a different associated strong lensing bias.
However, when considering only lenses with $\tein > 1''$, lens-selection and source-selection yield the same sample (see red histogram in \Fref{fig:lenspars}): in the limit of large Einstein radius, the two operations (identifying suitable lens galaxies and detecting a strongly lensed source) commute.

%In general, speaking of a lens- or source-selected sample is improper.
In general, there is no such thing as a purely lens- or source-selected sample. 
A lens-based search is always also source-based, because a detection of the lensed source is required to identify a lens.
A source-based search is also lens-based, because the lens properties determine important observational features of the lensed source, such as the number of images, the image separation and the total flux.
}

@@

{\bf
Our extended source analysis was based on a peak detection criterion to define strong lensing events, as explained in section \ref{ssec:lensdefext}.
However, many previous studies aimed at simulating populations of lenses adopted a magnification-based definition, in which a lens-source system is classified as a strong lens if the magnification of the source is larger than a minimum value \citep[see e.g.][]{Col15, Robertson++20}.
In order to check whether our results are robust to the choice of definition of a strong lensing event, we re-analysed the fiducial sample with this alternative definition. %repeated the analysis of the fiducial sample with extended sources. 
In particular, we defined as strong lenses all systems with a detected background source and $\mu_{\mathrm{tot}} > 3$, where $\mu_{\mathrm{tot}}$ is the total magnification (obtained by integrating the flux of all of the lensed images and dividing it by the intrinsic source flux).
The total number of strong lenses increased by $34\%$ with this new definition, mostly due to the inclusion of fainter sources with no detected multiple images.
\Fref{fig:minmubias} shows the strong lensing bias on a few selected quantities (stellar population synthesis mismatch parameter, projected dark matter mass within $5$~kpc, source magnitude and source half-light radius).
Besides the change in the source magnitude (third panel), we see very little differences in the bias with respect to the peak-based definition of the fiducial analysis. We then conclude that the exact definition of a strong lensing event does not have a big impact on the strong lensing bias of lens-related properties.
We stress out, however, that magnification is not a directly-observable quantity, and therefore we advise against applying magnification-based lens definitions to define lens samples, for the reasons discussed in section \ref{ssec:lensdefext}.
\begin{figure}
\includegraphics[width=\columnwidth]{minmu_bias.eps}
\caption{
Strong lensing bias in the case of a magnification-based lens definition (brown curves), compared to that of the fiducial analysis (blue curves).
First panel: median $\log{\asps}$.
Second panel: median $\mdmfive$ at $\log{\mstar}=11.5$ and $\log{\reff}=1.2$.
Third panel: median source magnitude.
Fourth panel: median half-light radius in the magnitude bin $24.8 < \msource < 25.2$.
}
\label{fig:minmubias}
\end{figure}
}

@@

\item Simulations with lensed quasars in place of extended sources showed that the amplitude of the strong lensing bias in the lens-related parameters has very little sensitivity on the details of the source population.
This result has positive implications for time-delay lensing studies: it means that information from a sample of galaxy-galaxy lenses can be used as a prior on the properties of a set of galaxy-quasar lenses, as long as the two samples are probing the same range in Einstein radius and lens observable properties.
{\bf
The associated bias on the inference of $H_0$ is on the order of $1\%$.
}
\item Samples of quad lenses are biased towards galaxies with larger ellipticity, which implies that their three-dimensional structure is also biased. This means that particular care must be taken when stellar dynamics measurements obtained on galaxy-galaxy lenses is used to inform the properties of quads.
\end{enumerate}

@@

{\bf
In the time-delay study of \citet{Bir++20}, lenses are described with a mass density profile given by \Eref{eq:mstprofile}.
\citet{Bir++20} defined the profile by first fitting a pure power-law mass model to the strong lensing data, thus constraining the Einstein radius and power-law index $\gamma$, and then applying a mass-sheet transformation of the following kind:
\begin{equation}\label{eq:mst}
\kappa(\boldsymbol\theta) \rightarrow \lmst\kappa(\boldsymbol\theta) + 1 - \lmst.
\end{equation}
%\citet{Bir++20} fixed the slope $\gamma$ of the power-law component to the value obtained by fitting a pure power-law mass model to the strong lensing data.
The mass-sheet transformation parameter $\lmst$ can be interpreted as a quantity describing a departure from a pure power-law profile.
One possible way of estimating $\lmst$ for our lenses is to fit a power-law to the simulated images and then optimize for the value of $\lmst$ that best matches the true density profile. That, however, would be too time-consuming. Instead, we emulate this process as follows.

\citet{Son18} showed that, when the main arc and the counter-image of an extended source are well resolved, their relative widths constrain the following combination of derivatives of the lens potential at the Einstein radius:
\begin{equation}\label{eq:invariant}
\frac{\psi'''}{1 - \psi''}.
\end{equation}
In other words, the above quantity can be measured robustly (i.e. in a model-independent way).
When fitting a pure power-law lens model to lensing data, then, the value of $\gamma$ that we obtain is the one that reproduces the true $\psi'''/(1-\psi'')$\footnote{Strictly speaking, this is only true in the limit of well-resolved arcs and nearly axisymmetric lenses}. This is given by
\begin{equation}
\gamma = 2 + \tein \frac{\psi'''}{1-\psi''}
\end{equation}
Hence, given a lens in our sample, we first define a pure power-law model with the same Einstein radius as the true one and with $\gamma$ given by the equation above. 
Then, given a pure power-law model, we find $\lmst$ such that the first three derivatives of the lens potential match the true values.
In particular, since a mass-sheet transformation changes the second derivative of the potential as
\begin{equation}
\psi'' \rightarrow \lmst\psi'' + 1 - \lmst,
\end{equation}
we set
\begin{equation}
\lmst = \frac{1 - \psi''_{\mathrm{PL}}}{1 - \psi''},
\end{equation}
where $\psi''_{\mathrm{PL}}$ is the value of $\psi''$ of the pure power-law lens model.
}

>>>>

  \abstract
  % context heading (optional)
  % {} leave it empty if necessary  
    {
Strong lenses are a biased subset of the general population of galaxies.
}
  % aims heading (mandatory)
   {
The goal of this work is to quantify how lens galaxies and lensed sources differ from their parent distribution, namely the strong lensing bias.
} 
   % methods heading (mandatory)
   {
We first studied how the strong lensing cross-section varies as a function of lens and source properties. 
Then, we simulated strong lensing surveys with data similar to that expected for Euclid and measured the strong lensing bias in different scenarios.
We focused particularly on two quantities: the stellar population synthesis mismatch parameter, $\asps$, defined as the ratio between the true stellar mass of a galaxy and the stellar mass obtained from photometry, and the central dark matter mass at fixed stellar mass and size.
}
% results heading (mandatory)
   {
Strong lens galaxies are biased towards larger stellar masses, smaller half-mass radii and larger dark matter masses.
The amplitude of the bias depends on the intrinsic scatter in the mass-related parameters of the galaxy population and on the completeness in Einstein radius of the lens sample.
For values of the scatter that are consistent with observed scaling relations and a minimum detectable Einstein radius of $0.5''$, the strong lensing bias in $\asps$ is $10\%$, while that in the central dark matter mass is $5\%$.
The bias has little dependence on the properties of the source population: samples of galaxy-galaxy lenses and galaxy-quasar lenses that probe the same Einstein radius distribution are biased in a very similar way.
}
  % conclusions heading (optional), leave it empty if necessary 
   {
Given current uncertainties, strong lensing observations can be used directly to improve our current knowledge of the inner structure of galaxies, without the need to correct for selection effects.
bf Time-delay measurements of $H_0$ from lensed quasars can take advantage of prior information obtained from galaxy-galaxy lenses with similar Einstein radius.
}
   \keywords{
             Gravitational lensing: strong 
               }

@@

Strong lenses, however, are a biased subset of the general population of galaxies and background sources.
In order for a lens-source pair to be included in a strong lensing survey, the following conditions must be met:
\begin{enumerate}
\item The lens must produce multiple images of the source
\item The multiple images must be detected and resolved
\item The system must be recognized as a lens by the survey.
\end{enumerate}
Condition 1 poses constraints on the mass distribution of the lens, as well as on the geometry of the system. 
First of all, not all lenses are capable of producing multiple images, but only those whose surface mass density $\Sigma(\boldsymbol\theta)$ is larger than the critical surface mass density for lensing $\Sigma_{\mathrm{cr}}$ at at least one position $\boldsymbol\theta$ \citep{SEF92}.
This condition excludes objects with a very diffuse mass distribution from the population of lenses.
Second, objects with a higher concentration of mass allow for a larger portion of the source plane to be multiply imaged, and therefore are over-represented in lens samples. Third, at fixed lens properties, sources that are located further away along the line of sight are more likely to be lensed.

@@

The cumulative effect of these conditions can be summarised as follows.
The probability distribution $\prsl$ of a sample of strong lenses from a given survey with selection criterion $S$ is given by \citep{Son22}:
\begin{equation}\label{eq:one}
\prsl(\psilens,\psisource|S) \propto \prlens(\psilens)\prsource(\psisource)\psel(\psilens,\psisource|S).
\end{equation}
In the equation above, $\psilens$ is the set of parameters describing the properties of foreground galaxies that are relevant for lensing, such as their redshift and mass distribution; $\psisource$ is the set of parameters describing background sources; $\prlens$ and $\prsource$ describe the parent distribution of foreground galaxies and background sources in the absence of lensing; and $\psel$ describes the probability of selecting a lens-source system with parameters $\psilens$ and $\psisource$ given the criterion $S$ used to define a lens.
This last factor takes into account both physical effects, that is whether a lens with parameters $\psilens$ produces a strongly lensed image of a source with parameters $\psisource$, and survey selection effects, that is whether such an image can be detected and recognised as a strong lens.

@@

Finally, we considered galaxy-quasar lenses. 
The strong lensing bias of lensed quasars was studied by \citep{C+C16}, in the context of power-law lens mass models with external convergence.
Here we addressed the question of how different are galaxy-quasar lenses from galaxy-galaxy lens samples. 
This point is relevant for time-delay cosmography studies, in which measurements of the time-delay between the multiple images of a strongly lensed quasar are used to constrain the expansion rate of the Universe. Galaxy-galaxy lenses can in principle be used to help break some of the model degeneracies affecting these measurements \citep{Bir++20,B+T21}, but any difference between the two lens classes can introduce biases, if not corrected for.
With this study we aim to quantify this bias.

@@

@@

The first four rows of \Fref{fig:quasarbias} show the strong lensing bias as a function of minimum Einstein radius in the following quantities: stellar population synthesis mismatch parameter, halo mass at fixed stellar mass, enclosed dark matter mass at fixed stellar mass and half-light radius, and axis ratio.
The biases of the population of lensed quasars (red curves) are very similar to that of lensed galaxies (blue curves), especially for values of $\theta_{\mathrm{Ein},min}<1''$, 
When considering only quad lenses, however, there are some differences, the most remarkable of which is the bias in the axis ratio: quad lenses tend to be on average galaxies with a higher ellipticity.
This is a well-known effect \citep[see e.g.][]{KKS97} which can be explained in the context of the results of section \ref{ssec:axisymmpoint}: the higher the ellipticity, the larger the area enclosed within the inner caustic, which is where a source needs to lie in order to produce four or more images.
Quad lenses tend to also have a slightly larger halo mass at fixed stellar mass, for $\theta_{\mathrm{Ein},min}<1''$.

@@

A measurement of the time delay between the images of a lensed quasar can be converted into an estimate of the expansion rate of the Universe, $H_0$. 
In general, though, the value of $H_0$ is degenerate with the mass distribution of the lens (and of matter along the line-of-sight), which can be difficult to constrain on an individual lens basis.
One possible strategy to break this denegeracy is to use information gathered from a larger set of lenses as a prior for the time-delay lenses.
This was the strategy adopted by \citet{Bir++20} in the most recent measurement of $H_0$ by the TDCOSMO collaboration: they combined a larger sample of extended source lenses with a smaller set of time-delay lenses.
In this section we want to quantify the possible bias on $H_0$ that can result from this approach.

The lens mass model used by \citet{Bir++20} consists in a power-law density profile modified by a mass-sheet transformation. The corresponding projected density profile is
\begin{equation}\label{eq:mstprofile}
\kappa(\theta) \propto \lmst\theta^{1-\gamma} + (1 - \lmst),
\end{equation}
where $\lmst$ is called the mass-sheet transformation parameter.
Strong lensing data can constrain very well the power-law index $\gamma$, but not at all $\lmst$.
Given the time delay, the inferred value of the Hubble constant $H_0$ scales linearly with $1/\lmst$.
\citet{Bir++20} measured the population distribution of $\lmst$ on the sample of extended source lenses, by combining strong lensing and stellar kinematics data, and then used it as a prior in the time-delay analysis.
Although the mass model used in our work is different from that of \Eref{eq:mstprofile}, we can estimate the bias associated with this approach by computing the value of $\lmst$ for each lens and then comparing the strong lensing bias on $\lmst$ of extended source and quasar lenses.
Appendix~\ref{sect:appendixc} explains how we obtained $\lmst$ for each lens. The strong lensing bias on the median value of $\lmst$ is shown in the fifth panel of \Fref{fig:quasarbias}\footnote{Since $\lmst$ is only defined for strong lenses, the median value of $\lmst$ of the general population of galaxies is not defined.}.

At fixed minimum Einstein radius, there is a small difference between the median $\lmst$ in extended source and quasar lenses, typically of order 1\%.
The associated bias on $H_0$ is, therefore, equally small.
The value of $\lmst$, however, increases with increasing minimum $\tein$. This implies that if the two samples probe different ranges in Einstein radius, the bias on $H_0$ can be larger than that.
The bias can also be larger if the properties of the general population of galaxies (as opposed to the strong lenses) are used to inform the structure of time-delay lenses \citep[see][]{C+C16}. This is simply because quasar lenses are on average more similar to galaxy-galaxy lenses than they are to the underlying galaxy population (the blue and red curves in \Fref{fig:quasarbias} are closer to each other than they are to the dashed horizontal lines).

@@

Strictly speaking, the results shown in \Sref{sect:results} apply only to samples of lenses and background sources with the same properties as our simulations.
Real lens samples can have more complex source distributions. For example, many existing lens-finding algorithms are tuned to identify sources with different colours from the lenses, introducing a selection that can modify the source distribution in redshift-magnitude space.
However, the comparison of section \ref{ssec:quasarbias} shows that, at fixed foreground galaxy population, the strong lensing bias for the sample of extended sources is indistinguishable from that of the lensed quasar population, at least for values of $\theta_{\mathrm{Ein,min}} < 1.0''$.
This result suggests that the dependence of the strong lensing bias on the details of the background source population is very weak, as long as no additional selection on the image configuration is applied (e.g. by selecting only quad lenses).
This is not a coincidence, but is a consequence of the fact that the dependence of the strong lensing cross-section on the lens parameters is a weak function of source magnitude\footnote{The cross-section itself is a strong function of source magnitude, but the trends between $\crosssect$ and the lens parameters are not.}, especially for magnitudes close to the detection limit (see section \ref{ssec:axisymmpoint}).
We can then conclude that the details of the properties of the background source population play a secondary role in determining the strong lensing bias.
A selection aimed at specific kinds of sources, such as blue star-forming galaxies or quasars, results primarily in different numbers of lenses being discovered, but does not change how biased the lenses are with respect to the general galaxy population.

@@

The simulations described in \Sref{sect:lenspop}, on which our experiment is based, are examples of so-called lens-selected surveys:
the lens samples are built by first defining a population of possible lens galaxies and then searching for strong lenses among them. Photometric lens searches are typically carried out in this way \citep[see e.g.][]{Gav++12,Son++18a,Pet++19}.
An alternative approach is to build a source-selected sample, by first searching for signatures of strong lensing within a population of possible background sources and then obtaining information on the lens galaxies \citep[see e.g.][]{Hez++13}.
In principle, these two approaches lead to different strong lensing biases. 
For example, from \Fref{fig:lenspars} we can see that the stellar mass distribution of the lenses extends below the minimum mass cut that we applied to the population of foregrounds galaxies: a source-based search that is able detect those lenses would result in a different sample and a different associated strong lensing bias.
However, when considering only lenses with $\tein > 1''$, lens-selection and source-selection yield the same sample (see red histogram in \Fref{fig:lenspars}): in the limit of large Einstein radius, the two operations (identifying suitable lens galaxies and detecting a strongly lensed source) commute.

In general, there is no such thing as a purely lens- or source-selected sample. 
A lens-based search is always also source-based, because a detection of the lensed source is required to identify a lens.
A source-based search is also lens-based, because the lens properties determine important observational features of the lensed source, such as the number of images, the image separation and the total flux.

@@

Our extended source analysis was based on a peak detection criterion to define strong lensing events, as explained in section \ref{ssec:lensdefext}.
However, many previous studies aimed at simulating populations of lenses adopted a magnification-based definition, in which a lens-source system is classified as a strong lens if the magnification of the source is larger than a minimum value \citep[see e.g.][]{Col15, Robertson++20}.
In order to check whether our results are robust to the choice of definition of a strong lensing event, we re-analysed the fiducial sample with this alternative definition. 
In particular, we defined as strong lenses all systems with a detected background source and $\mu_{\mathrm{tot}} > 3$, where $\mu_{\mathrm{tot}}$ is the total magnification (obtained by integrating the flux of all of the lensed images and dividing it by the intrinsic source flux).
The total number of strong lenses increased by $34\%$ with this new definition, mostly due to the inclusion of fainter sources with no detected multiple images.
\Fref{fig:minmulbias} and \Fref{fig:minmusbias} show the strong lensing bias on a series of lens- and source-related properties.
Besides the change in the source redshift and magnitude (first two panels in \Fref{fig:minmusbias}), we see very little differences in the bias with respect to the peak-based definition of the fiducial analysis. We then conclude that the exact definition of a strong lensing event does not have a big impact on the strong lensing bias of lens-related properties.
We stress out, however, that magnification is not a directly-observable quantity, and therefore we advise against applying magnification-based lens definitions to define lens samples, for the reasons discussed in section \ref{ssec:lensdefext}.
\begin{figure}
\includegraphics[width=\columnwidth]{minmu_lens_bias.eps}
\caption{
Lens strong lensing bias in the case of a magnification-based lens definition (brown curves), compared to that of the fiducial analysis (blue curves).
First panel: median $\log{\asps}$.
Second panel: median $\mdmfive$ at $\log{\mstar}=11.5$ and $\log{\reff}=1.2$.
Third panel: median $\gammadm$ at $\log{\mobs}=11.4$ and $\log{\reff}=1.2$.
Fourth panel: median axis ratio.
Fifth panel: median mass-sheet transformation parameter, defined in Appendix~\ref{sect:appendixc}.
In each panel, the dashed line indicates the value of the parent population.
}
\label{fig:minmulbias}
\end{figure}

\begin{figure}
\includegraphics[width=\columnwidth]{minmu_source_bias.eps}
\caption{
Source strong lensing bias in the case of a magnification-based lens definition (brown curves), compared to that of the fiducial analysis (blue curves). 
First panel: median source redshift.
Second panel: median source magnitude.
Third panel: median half-light radius in the magnitude bin $24.8 < \msource < 25.2$.
The dashed line indicates the median value of the population of sources detectable in the absence of lensing.
}
\label{fig:minmusbias}
\end{figure}
%__________________________________________________________________

@@

\item Simulations with lensed quasars in place of extended sources showed that the amplitude of the strong lensing bias in the lens-related parameters has very little sensitivity on the details of the source population.
This result has positive implications for time-delay lensing studies: it means that information from a sample of galaxy-galaxy lenses can be used as a prior on the properties of a set of galaxy-quasar lenses, as long as the two samples are probing the same range in Einstein radius and lens observable properties.
The associated bias on the inference of $H_0$ is on the order of $1\%$.
\item Samples of quad lenses are biased towards galaxies with larger ellipticity, which implies that their three-dimensional structure is also biased. This means that particular care must be taken when stellar dynamics measurements obtained on galaxy-galaxy lenses is used to inform the properties of quads.
\end{enumerate}

@@

In the time-delay study of \citet{Bir++20}, lenses are described with a mass density profile given by \Eref{eq:mstprofile}.
\citet{Bir++20} defined the profile by first fitting a pure power-law mass model to the strong lensing data, thus constraining the Einstein radius and power-law index $\gamma$, and then applying a mass-sheet transformation of the following kind:
\begin{equation}\label{eq:mst}
\kappa(\boldsymbol\theta) \rightarrow \lmst\kappa(\boldsymbol\theta) + 1 - \lmst.
\end{equation}
The mass-sheet transformation parameter $\lmst$ can be interpreted as a quantity describing a departure from a pure power-law profile.
One possible way of estimating $\lmst$ for our lenses is to fit a power-law to the simulated images and then optimize for the value of $\lmst$ that best matches the true density profile. That, however, would be too time-consuming. Instead, we emulate this process as follows.

\citet{Son18} showed that, when the main arc and the counter-image of an extended source are well resolved, their relative widths constrain the following combination of derivatives of the lens potential at the Einstein radius:
\begin{equation}\label{eq:invariant}
\frac{\psi'''}{1 - \psi''}.
\end{equation}
In other words, the above quantity can be measured robustly (i.e. in a model-independent way).
When fitting a pure power-law lens model to lensing data, then, the value of $\gamma$ that we obtain is the one that reproduces the true $\psi'''/(1-\psi'')$\footnote{Strictly speaking, this is only true in the limit of well-resolved arcs and nearly axisymmetric lenses}. This is given by
\begin{equation}
\gamma = 2 + \tein \frac{\psi'''}{1-\psi''}
\end{equation}
Hence, given a lens in our sample, we first define a pure power-law model with the same Einstein radius as the true one and with $\gamma$ given by the equation above. 
Then, given a pure power-law model, we find $\lmst$ such that the first three derivatives of the lens potential match the true values.
In particular, since a mass-sheet transformation changes the second derivative of the potential as
\begin{equation}
\psi'' \rightarrow \lmst\psi'' + 1 - \lmst,
\end{equation}
we set
\begin{equation}
\lmst = \frac{1 - \psi''_{\mathrm{PL}}}{1 - \psi''},
\end{equation}
where $\psi''_{\mathrm{PL}}$ is the value of $\psi''$ of the pure power-law lens model.
